Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures

### 2017-10-03 qnml.tex
subsection on regularity

\begin{proof}
We can write the terms on the left side of the equation as
\begin{eqnarray*}
P(D_{i, G_i} \ | \ \hat{\theta}_{i, G_i}) &=& \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}, \text{ and }\\
P(D_{G_i} \ | \ \hat{\theta}_{G_i} ) &=&  \prod_{j} \Le \frac{N_{ij}}{N}  \Ri^{N_{ij}}.
\end{eqnarray*}
Here, $N_{ijk}$ denotes the number of times we observe $X_i$ taking value $k$ when its parents are in $j^\text{th}$ configuration in our data matrix $D$. Also, $N_{ij} = \sum_k N_{ijk}$ (and $\sum_{k,j}N_{ijk} = N$ for all $i$).
Therefore,
\begin{align*}
\frac{P(D_{i, G_i} \ | \ \hat{\theta}_{i, G_i} )}{P(D_{G_i} \ | \ \hat{\theta}_{G_i} )} &= \frac{ \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}}{\prod_{j} \Le \frac{N_{ij}}{N}  \Ri^{N_{ij}}} \\
&= \frac{ \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}}{\prod_{j} \Le \frac{N_{ij}}{N}  \Ri^{\sum_k N_{ijk}}} \\
&= \frac{ \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}}{\prod_{j}\prod_{k} \Le \frac{N_{ij}}{N}  \Ri^{N_{ijk}}}\\
%% &= \prod_{j,k} \Le \frac{N_{ijk}}{N_{ij}} \Ri^{N_{ijk}} \\
&= P(D_i \ | \ \hat{\theta}_{i | G_i} ).
\end{align*}

@@

>>>>

\begin{proof}
We can write the terms on the left side of the equation as
\begin{eqnarray*}
P(D_{i, G_i} \ | \ \hat{\theta}_{i, G_i}) &=& \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}, \text{ and }\\
P(D_{G_i} \ | \ \hat{\theta}_{G_i} ) &=&  \prod_{j} \Le \frac{N_{ij}}{N}  \Ri^{N_{ij}}.
\end{eqnarray*}
Here, $N_{ijk}$ denotes the number of times we observe $X_i$ taking value $k$ when its parents are in $j^\text{th}$ configuration in our data matrix $D$. Also, $N_{ij} = \sum_k N_{ijk}$ (and $\sum_{k,j}N_{ijk} = N$ for all $i$).
Therefore,
\begin{align*}
\frac{P(D_{i, G_i} \ | \ \hat{\theta}_{i, G_i} )}{P(D_{G_i} \ | \ \hat{\theta}_{G_i} )} &= \frac{ \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}}{\prod_{j} \Le \frac{N_{ij}}{N}  \Ri^{N_{ij}}} \\
&= \frac{ \prod_{j,k} \Le \frac{N_{ijk}}{N}  \Ri^{N_{ijk}}}{\prod_{j}\prod_{k} \Le \frac{N_{ij}}{N}  \Ri^{N_{ijk}}}\\
%% &= \prod_{j,k} \Le \frac{N_{ijk}}{N_{ij}} \Ri^{N_{ijk}} \\
&= P(D_i \ | \ \hat{\theta}_{i | G_i} ).
\end{align*}

@@

\subsection{QNML IS REGULAR}

\cite{Suzuki2017} defines regularity for a scoring function $Q_n(X \mid Y)$ as follows:
\begin{definition}
Assume $H_N(X \mid Y') \leq H_N(X \mid Y)$, where $Y' \subset Y.$ We say that $Q_N(\cdot \mid \cdot)$ is regular if $Q_n(X \mid Y') \geq Q_N(X \mid Y)$.
\end{definition}
In the definition, $N$ denotes the sample size, $X$ is some random variable, $Y$ denotes the proposed parent set for $X$, and $H_N(\cdot \mid \cdot)$ refers to the empirical conditional entropy. \cite{Suzuki2017} shows that BDeu violates this principle and demonstrates that this causes the score to prefer more complex networks even though data do not support this. In \cite{Suzuki2017_2}, the authors also suggest that regular scores are computationally more efficient when applied with branch-and-bound type algorithms for Bayesian network structure learning.

By analyzing the penalty term of the qNML scoring function, one can prove the following statement:
\begin{theorem}
qNML score is regular.
\end{theorem}
\begin{proof}
The proof is given in the Appendix A in Supplementary Material.
\end{proof}
